Figures to insert in text

### 2018-04-15 Chapters/Chapter02.tex
Desiderata Table

    You know those animals that create those burrows underground, creatures like the prairie dogs?  There, underground, they build all sorts of passageways from here to where they want to go; and in the process they discover all that \emph{can't be passageway}.  What can't be passageway --- what can't be corridor --- will never be used as passageway, by definition.  And most of the world can't be passageway.  Think about the animal behavior: the beast won't give up! To me it's all very similar to what these systems seem to be doing... it seems they are molding a model of their problem; discovering what can't be passageway and finding some other route, very tortuous if need be...  You have these monstrous networks creating layers upon layers in which they differentiate and distribute the errors throughout all the nooks and cranies of, say, a domain like the game of Go.  Then, after some biblical number-crunching, it seems that there may be a very subtle way... however tortuous, however tight, however inprobable, however unconfortable.

@@

    But it feels like a mistake to say that that is thinking.  It's not.  It is to me a great model of someone getting so attuned into an area that they are able to simply flow through tasks; in the positive psychology sense of the word flow.  It's very powerful and very intelligent:  but it is a form of intelligence without thinking.  Intelligence that can't make leaps from one arena to another; Intelligence that can't work on Bongard problems.  So what I would fear as a dead-end in research would be to have decades of massive, massive number-crunching and machine learning of this type... in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figures \ref{fig:bongard-problems-71, fig:bongard-problems-74, fig:bongard-problems-77}; see \citet{bongard_recognition_1968, douglas1980godel, Linhares2000, foundalis_phaeaco:_2006, lupyan_difficulties_2013}).

    Cognitive Science is, to me, the study of plausible interaction between system 1 and system 2.  We can't continue to study, say, a silly, implausible, model of system 1 and a great model of system 2 (or vice-versa). We need both system 1 and system 2 to be plausible, in the psychological sense and in the neuroscientific sense.  AlphaGo, for instance, has this wonderful model of the game, reflecting many of the properties of system 1... but then falls back into silliness as soon as it turns to the other side of the problem:  it uses some montecarlo methods in very efficient fashion to go on to to the next probing of this wonderful system 1...

    Hofstadter has nailed the interactions and the content... analogies, fluid concepts, things seen through a different perspective, variations on a theme, etc.  Kanerva has nailed system 1, long-term-memory, and its  theoretical neuroscience; micro neuro-firings bringing a memory back to life.  Attaching something as beautiful as SDM to, say, some hyper-efficient centralized alpha-beta pruning algorithm is nothing short of heresy against all the gods of science for which the villain must be put to death without mercy. Seems like the only reasonable path going forward...

@@

@@

Ralph Waldo Emerson once said: \emph{do not go where the path may lead. Go, instead, where there is no path, and leave a trail.}  Professor Pentti Kanerva has left the trail.  It is my job to illuminate it and to pave it and to clear it; to try to deliver an easier pathway for the next generation.  Some essays completely shut the door close at the end; this one intends to leave it wide open. As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desideratum}, a prologue, a yearning for others to join me in imagining the shape of things to come.

>>>>

    You know those animals that create those burrows underground, creatures like the prairie dogs?  There, underground, they build all sorts of passageways from here to where they want to go; and in the process they discover all that \emph{can't be passageway}.  What can't be passageway --- what can't be corridor --- will never be used as passageway, by definition.  And most of the world can't be passageway.  Think about the animal behavior: endless dead-end nooks and cranies yet the beast won't give up! To me it's all very similar to what these systems seem to be doing... it seems they are molding a model of their problem; discovering what can't be passageway and finding some other route, very tortuous if need be...  You have these monstrous networks creating layers upon layers in which they differentiate and distribute the errors throughout all the nooks and cranies of, say, a domain like the game of Go.  Then, after some biblical number-crunching, it seems that there may be a very subtle way... however tortuous, however tight, however inprobable, however unconfortable.

@@

    But it feels like a mistake to say that that is thinking.  It's not.  It is to me a great model of someone getting so attuned into an area that they are able to simply flow through tasks; in the positive psychology sense of the word flow.  It's very powerful and very intelligent:  but it is a form of intelligence without thinking.  Intelligence that can't make leaps from one arena to another; Intelligence that can't work on Bongard problems.  So what I would fear as a dead-end in research would be to have decades of massive, massive number-crunching and machine learning of this type... in disregard to big questions such as the interaction of short-term memory with SDM and, again, tasks like Bongard problems (Figures \ref{fig:bongard-problems-71} \ref{fig:bongard-problems-74} \ref{fig:bongard-problems-77}; see \citet{bongard_recognition_1968, douglas1980godel, Linhares2000, foundalis_phaeaco:_2006, lupyan_difficulties_2013}).

    Cognitive Science has to study the plausible interaction between system 1 and system 2.  We can't continue to study, say, a silly, implausible, model of system 1 and a great model of system 2 (or vice-versa). We need both system 1 and system 2 to be plausible, in the psychological sense and in the neuroscientific sense.  AlphaGo, for instance, has this wonderful model of the game, reflecting many of the properties of system 1... but then falls back into silliness as soon as it turns to the other side of the problem:  it uses some montecarlo methods in very efficient fashion to go on to to the next probing of this wonderful system 1...

    Hofstadter has nailed the interactions and the content... analogies, fluid concepts, things seen through a different perspective, variations on a theme, etc.  Kanerva has nailed system 1, long-term-memory, and its  theoretical neuroscience; micro neuro-firings bringing a memory back to life.  Attaching something as beautiful as SDM to, say, some hyper-efficient centralized alpha-beta pruning algorithm is nothing short of heresy against all the gods of science for which the villain shall be put to death without clemency or mercy. Seems like the only reasonable path going forward...

@@

% Preview source code for paragraph 8

\begin{table}[ht]
\begin{tabular}{|p{4cm}|p{7cm}|}  %{|c|c|}
\hline
Project & Observation\tabularnewline
\hline
\hline
Source-code  & C, OpenCL, Python (NVIDIA CUDA, FPGAs desirable in the future)\tabularnewline
\hline
Step-by-step notebook studies & replication and dissemination of studies\tabularnewline
\hline
Documentation of framework & 100+ pages currently\tabularnewline
\hline
Open-Source Creative Commons Book & Under consideration (to be created from this thesis)\tabularnewline
\hline
Collection of SDM links, results, resources & modeled after `Awesome Machine Learning'\citep{misiti_awesome-machine-learning:_2018} and `Awesome Artificial Intelligence'\citep{lewis_awesome-artificial-intelligence:_2018} lists, community
managed\tabularnewline
\hline
Slides from the Jupyter Notebooks & online course, currently in planning phase\tabularnewline
\hline
Videos following the Course Slides  & online course, currently in planning phase\tabularnewline
\hline
\end{tabular}

\caption{Desiderata of open-source, creative-common deliverables, longer-term.  Some of these items are mature for wide dissemination, whereas others are in planning phase.  For an online SDM course, for instance, we might partner with a Ph.D. program in the Teaching of Mathematics and Physics, such as \citet{im/ufrj_programa_nodate}.}

\end{table}

@@

Ralph Waldo Emerson once said: \emph{do not go where the path may lead. Go, instead, where there is no path, and leave a trail.}  Professor Pentti Kanerva has left the trail.  It is my job to illuminate it and to pave it and to clear it; to try to deliver an easier pathway for the next generation.  Some essays completely shut the door close at the end; this one intends to leave it wide open. This is not a real conclusion.  As the reader might have noticed, this final section does not read as an analysis of the work done; it reads, instead, as a \emph{desideratum}, a prologue, a yearning for others to join me in imagining the shape of things to come.
